Think deep for reef resilience

### 2025-08-10 paper/paper_draft.qmd
Re-order and flesh out sections
Added an initial stub in supplementary to describe model parameterisation and sampling.

To apply exploratory scenario modelling, the approach employed a modelling framework
involving two separate systems: ADRIA (the Adaptive Dynamic Reef Interventin Algorithms
platform) and CoralBlox. ADRIA is a multi-criteria decision support platform designed to
help reef managers address climate change challenges in coral reef systems [@iwanaga2023].
The ADRIA platform provides functionality for scenario generation and analysis as well as
providing the technical infrastructure to couple environmental and ecosystem
representations.

CoralBlox is a computationally efficient coral population model. It simulates
functional/morphological groups of corals over its life cycle, competing for available space
over time [@ribeirodealmeida2024]. The coupled framework is referred to as ADRIA-CoralBlox.
The coral demographic representation and associated model implementation are further
detailed in the supplementary material.

@@

**\[TI and BG to flesh out\]**

For each GCM, ran 4096 reference scenarios ("counterfactuals")...

### Factors influencing bleaching mortality

Tropical cyclones and impact of wave activity to reefs were not included in this
investigation. We scale coral bleaching impacts by a coefficient determined by the median
depth of each reef location, following Baird et al. [@baird2018], reducing the proportion of
coral colonies on a reef impacted by high heat stress as depth increases.

## Timeseries analysis

ADRIA-CoralBlox was run for 4096 **sampled** scenarios for each of the five GCMs (20,480
total model runs). A median absolute coral cover timeseries for each reef was calculated for
each GCM. For timeseries clustering reef cover timeseries were scaled by the total area of
the reef to represent cover as a percentage of the total reef area. The proportion of each
functional group occupying the reef (group evenness) was also investigated as a metric for
timeseries analysis, but was found to not sufficiently capture the declined condition of
reefs suffering from very low absolute coral cover.

Given the uncertainties present, the proportion of reefs that display capacity to maintain
coral cover above a given $\beta_{0}$ threshold is examined. An indicative range of 10 to
20% is used (inclusive, with a step size of 1%). The range was informed by Perry et al.,
[-@perry2015] and Desbians et al., [-@desbiens2021] providing plausible indications of the
lower and upper bounds respectively. Robustness of the conclusions drawn to the uncertainty
in the precise value of $\beta_{0}$ for all reefs can then be assessed by using the
indicative range.

The relationship between the number of years above $\beta_{0}$ and median reef depth was
then assessed with Spearman's rank correlation. Spearman's correlation coefficient was
adopted as reef depth and connectivity strength distributions are expected to be non-normal
and highly skewed. Furthermore, reefs are naturally of different types and experience
different conditions, and so the monotonic trends (whether the relationship consistently
increase or decrease) rather than the exact quantative relationship is of interest.

@@

@@

@@

>>>>

@@

To apply exploratory scenario modelling, the approach employed a modelling framework
involving two separate systems: ADRIA (the Adaptive Dynamic Reef Interventin Algorithms
platform) and CoralBlox. ADRIA is a multi-criteria decision support platform designed to
help reef managers address climate change challenges in coral reef systems [@iwanaga2023].
The ADRIA platform provides functionality for scenario generation and analysis as well as
providing the technical infrastructure to couple environmental and ecosystem
representations. CoralBlox is a computationally efficient coral population model. It
simulates functional/morphological groups of corals over its life cycle, competing for
available space over time [@ribeirodealmeida2024]. The coupled framework is referred to as
ADRIA-CoralBlox.

The coral demographic representation and associated model implementation are further
detailed in the supplementary material. Of important note, however, is that the
representation of bleaching mortalities follow Bozec et al., [-@bozec2022] albeit with a
scaling coefficient to incorporate the amelioration effect that depth may have, following
Baird et al., [@baird2018]. The scaling reduces the proportion of coral colonies impacted by
high heat stress as depth increases. Tropical cyclones and impact of wave activity on reefs
were not included in this investigation.

ADRIA-CoralBlox was run for 4096 scenarios for each of the five GCMs (20,480 total model
runs). Here, a scenario represents a unique combination of assumed environmental (namely one
of the five DHW trajectories) and ecological factors including coral growth and mortality
rates, coral thermal tolerances. Scenarios were generated using the low-discrepancy Sobol'
sampling method [@sobol1967; @sobol1993] with Owen Scrambling [@owen1995], a random-Quasi
Monte Carlo approach which balances scenario exploration with improved convergence
properties (compared to pseudo-random sampling). A detailed comparative analysis of the
sampling methods can be found in [@kucherenko2015].

For the purpose of scenario assessment, the resulting timeseries for each model run were
scaled by the total area of the reef to represent cover as a percentage of the total
estimated reef area. The proportion of each functional group occupying the reef (group
evenness) was also investigated as a metric for timeseries analysis, but was found to not
sufficiently capture the declined condition of reefs suffering from very low absolute coral
cover **\[TI: Ben, is the sentence here really important to include?\]**. These timeseries
of coral cover were then clustered based on their temporal behaviour.

@@

Given the uncertainties present, the proportion of modelled reefs that display capacity to
maintain coral cover above a given $\beta_{0}$ threshold is examined. An indicative range of
10 to 20% is used (inclusive, with a step size of 1%). The range was informed by Perry et
al., [-@perry2015] and Desbians et al., [-@desbiens2021] providing plausible indications of
the lower and upper bounds respectively. Robustness of the conclusions drawn to the
uncertainty in the precise value of $\beta_{0}$ for all reefs can then be assessed by using
the indicative range.

The relationship between the number of years above $\beta_{0}$ and median reef depth was
then assessed with Spearman's rank correlation. Spearman's correlation coefficient was
adopted as reef depth and connectivity strength distributions are expected to be non-normal
and highly skewed. Furthermore, reefs are naturally of different types and experience
different conditions, and so the monotonic trends (whether the relationship consistently
increase or decrease) rather than the exact quantative relationship was of interest for this
analysis.

@@

### Scenario generation

Scenarios were sampled from the range of environmental and ecological factors using the
low-discrepancy Sobol' sampling method [@sobol1967; @sobol1993] with Owen scrambling
[@owen1995]. A base of 2 and padding of 32 is used for the Owen scrambling. Ecological
factors were sampled using a +/- 10% range relative to their nominal (assumed best guess)
value.

@@

**Bleaching mortality**
